---
title: 'Class 11: Scraping Multiple Webpages'
date: "2026-09-30"
output: 
  html_document: 
    toc: true
    toc_float: true
    number_sections: true
    code_download: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Learning Objectives

- Enable me to scrape multiple webpages(from the same site), using a 'for' loop type approach.

- Enable me to scrape multiple webpages(from the same site), using the [purrr package](https://purrr.tidyverse.org/), which is a functional programming approach. 

*note:* In this process, we will create our own function

## Re-examining the bulleting example

Let us try to expand on the example from last class to include the following:

[ACC](https://bulletin.miamioh.edu/courses-instruction/acc/),
[BUS](https://bulletin.miamioh.edu/courses-instruction/bus/),
[ISA](https://bulletin.miamioh.edu/courses-instruction/isa/), and 
[MKT](https://bulletin.miamioh.edu/courses-instruction/mkt/)


**Comment:** Obviously, you can repeat your code few times with chenging the URL. This will work for a few links but would be quite tedious for large amounts of scraping. We will follow the **DRY Principle**, which stands for **Don't Repeat Yourself**.

```{r bulletin}
dept_abbrv = c('acc', 'bus', 'isa', 'mkt')


courses_df = data.frame() #Initializing a container to store stuff in from the loop
for (abbrv in dept_abbrv) {
  cat(rep("-", 50), "\n")
  print(abbrv)
  bulletin_url = paste0(
    'https://bulletin.miamioh.edu/courses-instruction/', abbrv, '/'
  )
  
  # Q. Why did fadel make this an R object instead of a string like last class?
  # [a] we did this so we can reuse the rest of our code with only changing the URL
  webpage = rvest::read_html(bulletin_url)
  
  webpage |> rvest::html_elements("strong") |> rvest::html_text2() -> titles
  webpage |> rvest::html_elements("p.courseblockdesc") |> rvest::html_text2() -> descs
  
  df = data.frame(title = titles, desc = descs)
  print(df$title[1:2])
  courses_df = rbind(courses_df, df)
}

```

## 'purrr' approach for the same problem

this is a functional programming approach, and we will write our own function. In R, we use a function called 'function' to create a function. As opposed to python, R functions return a single object.

```{r purr}

dept_abbrv = c('acc', 'bus', 'isa', 'mkt')
scrape_mu_bulleting = function(abbrv){
   bulletin_url = paste0(
    'https://bulletin.miamioh.edu/courses-instruction/', abbrv, '/'
  )
  
  # Q. Why did fadel make this an R object instead of a string like last class?
  # [a] we did this so we can reuse the rest of our code with only changing the URL
  webpage = rvest::read_html(bulletin_url)
  
  webpage |> rvest::html_elements("strong") |> rvest::html_text2() -> titles
  webpage |> rvest::html_elements("p.courseblockdesc") |> rvest::html_text2() -> descs
  
  df = data.frame(title = titles, desc = descs)
  return(df)
}

test = scrape_mu_bulleting(abbrv ='isa')

# to use functions from purr, we will use a map_*(), or map2_*(), or a pmap_*()
# map uses a function that takes a single input(or you are only changing a single input)
# map2 uses a function that takes two inputs
# pmap >=3


courses_df_from_purr = purrr::map_df(.x = dept_abbrv, .f = scrape_mu_bulleting)
```


## scrape data from the website planecrashinfo 

the website that we want to scrape is called: <https://www.planecrashinfo.com/> and it contains a database of plane crashes by year. 
```{r}
year = 2025

crash_url = paste0('https://www.planecrashinfo.com/', year, '/', year, )
```


