---
title: "Midwest Airbnb Market Analysis"
author: "Ethan Williams"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float: true
    number_sections: true
    code_download: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

This report looks at Airbnb listings in Chicago, Columbus, and the Twin Cities.
The data comes from three tables: one row per listing, one row per listing per
month of upcoming availability, and one row per listing per month of reviews.

# Reading the data

```{r read-data}
con = DBI::dbConnect(RSQLite::SQLite(), "../data/midwest_airbnb_extended.db")

listings     = DBI::dbReadTable(con, "listings") |> tibble::as_tibble()
availability = DBI::dbReadTable(con, "availability_monthly") |> tibble::as_tibble()
reviews      = DBI::dbReadTable(con, "reviews_monthly") |> tibble::as_tibble()

DBI::dbDisconnect(con)

dplyr::glimpse(listings)
```

# Task 1: Who owns the market

First I count how many listings each host has. Then I attach that count to every
listing, so each listing knows how big its host is.

```{r task-1}
host_counts = listings |>
  dplyr::count(host_id, name = "host_listings")

big_hosts = host_counts |>
  dplyr::filter(host_listings >= 10)

nrow(big_hosts)   
```

```{r task-1b}
listings_with_host_size = listings |>
  dplyr::left_join(host_counts, by = "host_id") |>
  dplyr::mutate(is_big_host = host_listings >= 10)

listings_with_host_size |>
  dplyr::group_by(city) |>
  dplyr::summarize(pct_big_host = round(100 * mean(is_big_host), 1))   
```

# Task 2: Value for a family of four

A family of four wants a whole place, room for at least four people, and a
track record. I narrow to those listings, work out the price per person, and
then ask which neighbourhood is cheapest.

```{r task-2}
family_listings = listings |>
  dplyr::filter(
    room_type == "Entire home/apt",
    accommodates >= 4,
    number_of_reviews >= 10,
    review_scores_rating >= 4.8
  ) |>
  dplyr::mutate(price_per_person = price / accommodates)

family_by_neighbourhood = family_listings |>
  dplyr::group_by(city, neighbourhood) |>
  dplyr::summarize(
    n_listings = dplyr::n(),
    median_ppp = median(price_per_person, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::filter(n_listings >= 20)

family_by_neighbourhood |>
  dplyr::filter(city == "Columbus") |>
  dplyr::arrange(median_ppp)   # A3 = the first row
```

# Task 3: What earns

Three medians at once, for every combination of city and room type.
`dplyr::across()` saves writing the same `median()` line three times.

```{r task-3}
by_city_room = listings |>
  dplyr::group_by(city, room_type) |>
  dplyr::summarize(
    dplyr::across(
      c(price, reviews_per_month, estimated_revenue_l365d),
      \(x) median(x, na.rm = TRUE)
    ),
    .groups = "drop"
  )

by_city_room |>
  dplyr::filter(room_type == "Entire home/apt")   # A4 = Twin Cities row
```

# Task 4: Concentration

Inside each city I rank the listings from the biggest earner down, flag the top
tenth, and see how much of the city's revenue those few listings bring in.

```{r task-4}
ranked_listings = listings |>
  dplyr::group_by(city) |>
  dplyr::arrange(dplyr::desc(estimated_revenue_l365d), .by_group = TRUE) |>
  dplyr::mutate(
    rank   = dplyr::row_number(),
    cutoff = ceiling(0.10 * dplyr::n()),
    in_top_10 = rank <= cutoff
  )

ranked_listings |>
  dplyr::summarize(
    top10_revenue_pct = round(
      100 * sum(estimated_revenue_l365d[in_top_10]) / sum(estimated_revenue_l365d), 1
    ),
    .groups = "drop"
  )   # A5 = Columbus row
```

# Task 5: The year ahead

October 2026 availability, whole homes only. If fewer than half the nights are
still open, I call the listing mostly taken.

```{r task-5}
entire_homes = listings |>
  dplyr::filter(room_type == "Entire home/apt") |>
  dplyr::select(id, city)

october = availability |>
  dplyr::filter(month == "2026-10") |>
  dplyr::inner_join(entire_homes, by = "id") |>
  dplyr::mutate(mostly_taken = nights_available < 0.5 * nights_in_month)

october |>
  dplyr::group_by(city) |>
  dplyr::summarize(pct_mostly_taken = round(100 * mean(mostly_taken), 1)) |>
  dplyr::arrange(pct_mostly_taken)   # A6 and A7 = the first row
```

# Task 6: Demand by season

Reviews stand in for completed stays, so counting reviews by month shows when
each city is busy. I keep the twelve months from July 2025 to June 2026.

```{r task-6}
listing_cities = listings |>
  dplyr::select(id, city)

reviews_by_month = reviews |>
  dplyr::inner_join(listing_cities, by = "id") |>
  dplyr::filter(month >= "2025-07", month <= "2026-06") |>
  dplyr::group_by(city, month) |>
  dplyr::summarize(total_reviews = sum(reviews), .groups = "drop")

twin_cities = reviews_by_month |>
  dplyr::filter(city == "Twin Cities") |>
  dplyr::arrange(dplyr::desc(total_reviews))

twin_cities   # A8 = the first row

round(max(twin_cities$total_reviews) / min(twin_cities$total_reviews), 2)   # A9
```

# Task 7: The Superhost premium

Do Superhosts charge more than everyone else? I find the median price for each
group in each city, then split the two groups apart and subtract.

```{r task-7}
superhost_summary = listings |>
  dplyr::filter(
    room_type == "Entire home/apt",
    host_is_superhost %in% c("t", "f")
  ) |>
  dplyr::group_by(city, host_is_superhost) |>
  dplyr::summarize(
    n = dplyr::n(),
    median_price  = median(price, na.rm = TRUE),
    median_rating = median(review_scores_rating, na.rm = TRUE),
    .groups = "drop"
  )

superhost_summary
```

```{r task-7b}
superhost_prices = superhost_summary |>
  dplyr::filter(host_is_superhost == "t") |>
  dplyr::select(city, superhost_price = median_price)

other_prices = superhost_summary |>
  dplyr::filter(host_is_superhost == "f") |>
  dplyr::select(city, other_price = median_price)

superhost_prices |>
  dplyr::inner_join(other_prices, by = "city") |>
  dplyr::mutate(premium = superhost_price - other_price)   # B1
```

# Task 8: What the join lost

```{r task-8}
missing_calendar = listings |>
  dplyr::anti_join(availability, by = "id")

nrow(missing_calendar)   # A10

missing_calendar |>
  dplyr::count(city, sort = TRUE)
```
